---
title: "Model free evidence bachelor thesis"
author: "Javier Torralba"
date: "2023-04-25"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

# Model free evidence for my thesis
## Load the libraries
```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(lubridate)
library(viridis)
```
## Loading data that will be used
```{r}
all_data_jt1_date_read_and_isbn <- read_csv("D:/Research assistant/Thesis/Data wrangling for Thesis/Preparation of datasets/all_data_jt1_date_read_and_isbn.csv")
```
## Preparing graph data
```{r}
stringency_index_graph <-   all_data_jt1_date_read_and_isbn %>% 
  select(date_read, stringency_index) %>% na.omit() %>%
mutate(week_start = floor_date(date_read, unit = "week"))  %>%
  group_by(week_start) %>%
  summarize(avg_stringency_index = mean(stringency_index))

stringency_index_graph <- stringency_index_graph %>%
  filter(week_start >= as.Date("2020-01-01"))
  
```

Stringency index graph
```{r}
stringency_index_plot <- ggplot() + 
  geom_line(data = stringency_index_graph, aes(x = week_start, y = avg_stringency_index), size = 0.8) + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red", linetype = "dashed") +
  labs(x = "",
       y = "Average stringency index",
       title = "Figure 3: Development of stringency index") +
    scale_x_date(date_labels = "%b-%Y", 
               breaks = seq(as.Date("2020-01-01"), as.Date("2022-01-01"), by = "2 month"), 
               limits = c(as.Date("2020-01-01"), as.Date("2022-01-01")),
               name = "") +
  
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 13, hjust = 0.5)) +
  ggtitle("Figure 3: Development of stringency index")

ggsave("stringency_index_graph.png", plot = stringency_index_plot, height = 4, width = 6, dpi = 300)
```

## Average user rating graph
Preparing data
```{r}
user_rating_graph <-   all_data_jt1_date_read_and_isbn %>% 
  select(date_read, user_rating) %>% 
  na.omit() %>%
mutate(week_start = floor_date(date_read, unit = "week"))  %>%
  group_by(week_start) %>%
  summarize(avg_user_rating = mean(user_rating)) %>%
    filter(week_start >= as.Date("2018-01-01"))
```


Typical boring graph 
```{r}
user_rating_plot <- ggplot() + 
  geom_line(data = user_rating_graph, aes(x = week_start, y = avg_user_rating), size = 0.28) + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red", linetype = "dashed") +
  labs(x = "",
       y = "Average rating (out of 5)",
       title = "Development of average user rating over time") +
      scale_x_date(date_labels = "%b-%Y", 
               breaks = seq(as.Date("2018-01-01"), as.Date("2022-01-01"), by = "3 month"), 
               limits = c(as.Date("2018-01-01"), as.Date("2022-01-01")),
               name = "") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 13, hjust = 0.5)) +
  ggtitle("Figure 4: Development of average user rating")
  
ggsave("user_rating_graph_UPDATED.png", plot = user_rating_plot, height = 4, width = 6, dpi = 300)
```

## Average reading time per day graph
```{r}
reading_time_days_graph <-   all_data_jt1_date_read_and_isbn %>% 
  select(date_read, read_time_days) %>%
  mutate(read_time_days = ifelse(read_time_days > quantile(read_time_days, 0.98, na.rm = TRUE), NA, read_time_days)) %>%
  na.omit() %>%
mutate(week_start = floor_date(date_read, unit = "week"))  %>%
  group_by(week_start) %>%
  summarize(avg_read_time_days = mean(read_time_days)) %>%
    filter(week_start >= as.Date("2018-01-01"))
```


Typical boring graph 
```{r}
reading_time_plot <- ggplot() + 
  geom_line(data = reading_time_days_graph, aes(x = week_start, y = avg_read_time_days), size = 0.4) + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red", linetype = "dashed") +
  labs(x = "",
       y = "Days",
       title = "Figure 5: Average number of days to read a book") +
      scale_x_date(date_labels = "%b-%Y", 
               breaks = seq(as.Date("2018-01-01"), as.Date("2022-01-01"), by = "3 month"), 
               limits = c(as.Date("2018-01-01"), as.Date("2022-01-01")),
               name = "") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 13, hjust = 0.5)) +
  ggtitle("Figure 5: Average number of days to read a book")
  
ggsave("reading_time_graph_UPDATED.png", plot = reading_time_plot, height = 4, width = 6, dpi = 300)

```
Adapted average number of reading days per graph
```{r}
reading_time_days_graph_adapted <-   all_data_jt1_date_read_and_isbn %>% 
  select(date_read, read_time_days) %>% 
  filter(read_time_days <= 22) %>%
  na.omit() %>%
mutate(week_start = floor_date(date_read, unit = "week"))  %>%
  group_by(week_start) %>%
  summarize(avg_read_time_days = mean(read_time_days)) %>%
      filter(week_start >= as.Date("2018-01-01"))
```

```{r}
ggplot() + 
  geom_line(data = reading_time_days_graph_adapted, aes(x = week_start, y = avg_read_time_days), size = 0.4) + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red", linetype = "dashed") +
  labs(x = "",
       y = "Days",
       title = "Figure 4 A: Filtered version for average number of days to read a book") +
      scale_x_date(date_labels = "%b-%Y", 
               breaks = seq(as.Date("2018-01-01"), as.Date("2022-01-01"), by = "6 month"), 
               limits = c(as.Date("2015-01-01"), as.Date("2022-01-01")),
               name = "") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 13, hjust = 0.5)) +
  ggtitle("Figure 4 A: Filtered version for average number of days to read a book")

ggsave("user_reading_time_per_day_adapted.png", plot = reading_time_days_graph_adapted_plot, height = 4, width = 6, dpi = 300)
```


## Average age of books graph
```{r}
average_age_of_book <-   all_data_jt1_date_read_and_isbn %>% 
  select(date_read, age_of_book) %>% 
  mutate(age_of_book = ifelse(age_of_book > quantile(age_of_book, 0.98, na.rm = TRUE), NA, age_of_book)) %>%
  na.omit() %>%
mutate(week_start = floor_date(date_read, unit = "week"))  %>%
  group_by(week_start) %>%
  summarize(avg_age_of_book = mean(age_of_book)) %>%
        filter(week_start >= as.Date("2018-01-01"))
```

Typical boring graph 
```{r}
 average_age_of_book_plot <- ggplot() + 
  geom_line(data = average_age_of_book, aes(x = week_start, y = avg_age_of_book), size = 0.3) + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red", linetype = "dashed") +
  labs(x = "",
       y = "Age of book (in years)",
       title = "Figure 6: Development of average age of books read") +
      scale_x_date(date_labels = "%b-%Y", 
               breaks = seq(as.Date("2018-01-01"), as.Date("2022-01-01"), by = "3 month"), 
               limits = c(as.Date("2018-01-01"), as.Date("2022-01-01")),
               name = "") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 13, hjust = 0.5))

ggsave("average_age_of_book_UPDATED.png", plot = average_age_of_book_plot, height = 4, width = 6, dpi = 300)
  
```
## Total number of pages
```{r}
number_of_pages_graph <- all_data_jt1_date_read_and_isbn %>%
  select(num_pages, date_read) %>% 
  na.omit() %>%
  mutate(week_start = floor_date(date_read, unit = "week"))  %>%
  group_by(week_start) %>%
  summarize(number_of_pages_read = sum(num_pages)) %>%
  mutate(total_pages_read = ifelse(number_of_pages_read > quantile(number_of_pages_read, 0.98, na.rm = TRUE), NA, number_of_pages_read)) %>%
  filter(week_start >= as.Date("2018-01-01"))
```

```{r}
number_of_pages_plot <- ggplot() + 
  geom_line(data = number_of_pages_graph, aes(x = week_start, y = number_of_pages_read), size = 0.3) + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red", linetype = "dashed") +
  labs(x = "",
       y = "Number of pages",
       title = "Figure 7: Total number of pages read in a week") +
      scale_x_date(date_labels = "%b-%Y", 
               breaks = seq(as.Date("2018-01-01"), as.Date("2022-01-01"), by = "3 month"), 
               limits = c(as.Date("2018-01-01"), as.Date("2022-01-01")),
               name = "") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 13, hjust = 0.5))

ggsave("number_of_pages_graph.png", plot = number_of_pages_plot, height = 4, width = 6, dpi = 300)
```

## Total number of books
```{r}
number_of_books_graph <- all_data_jt1_date_read_and_isbn %>%
  select(user_id, date_read) %>% 
  na.omit() %>%
  mutate(week_start = floor_date(date_read, unit = "week"), number_of_books = 1) %>%
  group_by(week_start) %>%
  summarize(sum_of_books = sum(number_of_books)) %>%
  filter(week_start >= as.Date("2018-01-01"))
```

```{r}
number_of_books_plot <- ggplot() + 
  geom_line(data = number_of_books_graph, aes(x = week_start, y = sum_of_books), size = 0.3) + 
  geom_vline(xintercept = as.numeric(as.Date("2020-03-01")), color = "red", linetype = "dashed") +
  labs(x = "",
       y = "Number of books",
       title = "Figure 8: Total number of books read in a week") +
      scale_x_date(date_labels = "%b-%Y", 
               breaks = seq(as.Date("2018-01-01"), as.Date("2022-01-01"), by = "3 month"), 
               limits = c(as.Date("2018-01-01"), as.Date("2022-01-01")),
               name = "") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 13, hjust = 0.5))

ggsave("number_of_books_graph_UPDATED.png", plot = number_of_books_plot, height = 4, width = 6, dpi = 300)
```













From an old document
```{r}
# Create a new column for year and month
users_graph$year_month <- format(users_graph$week_start, "%Y-%m")

# Create a new column for year
users_graph$year <- format(users_graph$week_start, "%Y")

# Find the first and last dates for each year
year_range <- users_graph %>%
  group_by(year) %>%
  summarize(min_date = min(week_start), max_date = max(week_start))

# Calculate the number of days between the first and last date for each year
year_range$year_duration <- as.numeric(difftime(year_range$max_date, year_range$min_date, units = "days"))

# Calculate the center point for each year's data
year_range$year_center <- year_range$min_date + year_range$year_duration / 2

# Create the plot
ggplot(users_graph, aes(x = week_start, y = avg_user_rating, color = year, group = year)) +
  geom_line(position = position_dodge(width = 0.5)) +
  scale_x_date(limits = c(min(year_range$year_center) - 180, max(year_range$year_center) + 180), 
               breaks = year_range$year_center, 
               labels = year_range$year, 
               date_labels = "%b") +
  labs(title = "Average User Rating by Month and Year",
       x = "Month",
       y = "Average User Rating",
       color = "Year")
```


Making the graph
```{r}
ggplot(data = users_graph, aes(x = week_start)) +
  geom_line(aes(y = avg_user_rating), size = 0.8) +
  geom_line(aes(y = avg_stringency_index/25), size = 0.8) +
  scale_y_continuous(name = "Average User Rating",
                     sec.axis = sec_axis(~.*25, name = "Average Stringency Index")) +
  labs(title = "Average User Rating and Stringency Index Over Time",
       x = "Two-Week Intervals", y = "Average User Rating",
       color = "Metric")


```

Removing data
```{r}
rm(users_graph)
```

## On reading time per days
Preparing data for graph
```{r}
read_time_days_graph <-   all_data_jt1_date_read_and_isbn %>% select(date_read, read_time_days, stringency_index) %>% na.omit() %>%
mutate(week_start = floor_date(date_read, unit = "week"))  %>%
  group_by(week_start) %>%
  summarize(avg_stringency_index = mean(stringency_index),
            avg_read_time_days = mean(read_time_days))
```

Making the graph
```{r}
ggplot(data = read_time_days_graph, aes(x = week_start)) +
  geom_line(aes(y = avg_read_time_days), size = 0.8) +
  geom_line(aes(y = avg_stringency_index/3), size = 0.8) +
  scale_y_continuous(name = "Average reading time per days",
                     sec.axis = sec_axis(~.*3, name = "Average Stringency Index")) 

```

Removing data
```{r}

```

## On Age of a book
Preparing data for graph
```{r}
age_of_book_graph <-   all_data_jt1_date_read_and_isbn %>% select(date_read, age_of_book, stringency_index) %>% na.omit() %>%
mutate(week_start = floor_date(date_read, unit = "14 days"))  %>%
  group_by(week_start) %>%
  summarize(avg_stringency_index = mean(stringency_index),
            avg_age_of_book = mean(age_of_book))
```

Making the graph
```{r}
ggplot(data = age_of_book_graph, aes(x = week_start)) +
  geom_line(aes(y = avg_age_of_book), size = 0.8) +
  geom_line(aes(y = avg_stringency_index/3), size = 0.8) +
  scale_y_continuous(
                     sec.axis = sec_axis(~.*3, name = "Average Stringency Index")) 
```

Removing data
```{r}

```

## On Total number of books read
Preparing data for graph
```{r}
number_of_books_graph <- all_data_jt1_date_read_and_isbn %>%
  select(user_id, date_read, stringency_index) %>% 
  na.omit() %>%
  mutate(bi_week_start = floor_date(date_read, unit = "week")) %>%
  group_by(bi_week_start) %>%
  summarize(count = n(), avg_stringency_index = mean(stringency_index))
```

Making the graph
```{r}
ggplot(data = number_of_books_graph, aes(x = bi_week_start)) +
  geom_line(aes(y = count), size = 0.6)
```

Removing data
```{r}

```

##On total number of pages read
Preparing data for graph
```{r}
number_of_pages_graph <- all_data_jt1_date_read_and_isbn %>%
  select(num_pages, date_read, stringency_index) %>% 
  na.omit() %>%
  mutate(bi_week_start = floor_date(date_read, unit = "week")) %>%
  group_by(bi_week_start) %>%
  summarize(count = sum(num_pages), avg_stringency_index = mean(stringency_index))
```

Making the graph
```{r}
ggplot(data = number_of_pages_graph, aes(x = bi_week_start)) +
  geom_line(aes(y = count), size = 0.6)
```

Removing data
```{r}

```
