---
title: "test_sprint_coverage_visualizations"
author: "Javier Torralba"
date: "2023-06-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  stars, # spatiotemporal data handling
  raster, # raster data handling
  terra, # raster data handling
  sf, # vector data handling
  dplyr, # data wrangling
  stringr, # string manipulation
  lubridate, # dates handling
  data.table, # data wrangling
  patchwork, # arranging figures
  tigris, # county border
  colorspace, # color scale
  viridis, # arranging figures
  tidyr, # reshape
  ggspatial, # north arrow and scale bar
  ggplot2, # make maps
  foreign,
  readr,
  maps
)
```

```{r}
us_states <- map_data("state")
```

What does each number of the files represent? (only for 2018 visualizations, numbering changes after 2019)

80  – WCDMA/UMTS/HSPA: 3G
81  – HSPA+: 3.5G
82  – EVDO/EVDO Rev A: 3G
83  – LTE: 4G
84  – WiMAX: 4G (can also support fixed wireless access)
85  – CDMA: 2G/3G (CDMA2000 can be considered 3G)
86  – GSM: 2G/2.5G
87  – Analog: 1G/Analog (legacy)
88  – Other: Varies (can include different generations or specific technologies)

This information can be found in the FCC Form 477 ("old477instructions.pdf"), table 3

Loading in data
```{r}
sprint_3G_2018_06 <- st_read(dsn = "F477_2018_06_Broadband_Sprint_82", layer = "Sprint_82")
sprint_4G_LTE_2018_06 <- st_read(dsn = "F477_2018_06_Broadband_Sprint_83", layer = "Sprint_83")
```


Data visualizations
```{r}
sprint_3G_2018_06_map <- ggplot() +
  geom_sf(data = sprint_3G_2018_06) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "Sprint 3G Coverage (EVDO/EVDO Rev A) June 2018") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("sprint_3G_2018_06_map.pdf", plot = sprint_3G_2018_06_map, height = 4, width = 6, dpi = 100)
```
```{r}
rm(sprint_3G_2018_06, sprint_3G_2018_06_map)
```

```{r}
sprint_4G_LTE_2018_06_map <- ggplot() +
  geom_sf(data = sprint_4G_LTE_2018_06) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "Sprint 4G LTE June 2018") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("sprint_4G_LTE_2018_06_map.pdf", plot = sprint_4G_LTE_2018_06_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(sprint_4G_LTE_2018_06, sprint_4G_LTE_2018_06_map)
```


```{r}
T_Mobile_3G_2018_06 <- st_read(dsn = "F477_2018_06_Broadband_T_Mobile_80", layer = "T_Mobile_80")
T_Mobile_3.5G_2018_06 <- st_read(dsn = "F477_2018_06_Broadband_T_Mobile_81", layer = "T_Mobile_81")
T_Mobile_4G_LTE_2018_06 <- st_read(dsn = "F477_2018_06_Broadband_T_Mobile_83", layer = "T_Mobile_83")
```
```{r}
T_Mobile_3G_2018_06_map <- ggplot() +
  geom_sf(data = T_Mobile_3G_2018_06) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 3G (WCDMA/UMTS/HSPA) June 2018") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_3G_2018_06_map.pdf", plot = T_Mobile_3G_2018_06_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(T_Mobile_3G_2018_06, T_Mobile_3G_2018_06_map)
```

```{r}
T_Mobile_3.5G_2018_06_map <- ggplot() +
  geom_sf(data = T_Mobile_3.5G_2018_06) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 3.5G (HSPA+) June 2018") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_3.5G_2018_06_map.pdf", plot = T_Mobile_3.5G_2018_06_map, height = 4, width = 6, dpi = "screen")
```

```{r}
rm(T_Mobile_3.5G_2018_06, T_Mobile_3.5G_2018_06_map)
```

```{r}
T_Mobile_4G_LTE_2018_06_map <- ggplot() +
  geom_sf(data = T_Mobile_4G_LTE_2018_06) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 4G LTE June 2018") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_4G_LTE_2018_06_map.pdf", plot = T_Mobile_4G_LTE_2018_06_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(T_Mobile_4G_LTE_2018_06, T_Mobile_4G_LTE_2018_06_map)
```

```{r}
Sprint_4G_LTE_2019_12 <- st_read(dsn = "F477_2019_12_Broadband_Sprint_83_268336", layer = "Sprint_LTE")
Sprint_5G_2019_12 <- st_read(dsn = "F477_2019_12_Broadband_Sprint_84_266740", layer = "Sprint_5G")
Sprint_3G_2019_12 <- st_read(dsn = "F477_2019_12_Broadband_Sprint_85_268203", layer = "Sprint_EVDO") 
```

```{r}
Sprint_5G_2019_12_map <- ggplot() +
  geom_sf(data = Sprint_5G_2019_12) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "Sprint 5G December 2019") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("Sprint_5G_2019_12_map.pdf", plot = Sprint_5G_2019_12_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(Sprint_5G_2019_12_map, Sprint_5G_2019_12)
```



```{r}
Sprint_4G_LTE_2019_12_map <- ggplot() +
  geom_sf(data = Sprint_4G_LTE_2019_12) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "Sprint 4G LTE December 2019") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("Sprint_4G_LTE_2019_12_map.pdf", plot = Sprint_4G_LTE_2019_12_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(Sprint_4G_LTE_2019_12, Sprint_4G_LTE_2019_12_map)
```


```{r}
Sprint_3G_2019_12_map <- ggplot() +
  geom_sf(data = Sprint_3G_2019_12) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "Sprint 3G December 2019") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("Sprint_3G_2019_12_map.pdf", plot = Sprint_3G_2019_12_map, height = 4, width = 6, dpi = 72)
```


```{r}
rm(Sprint_3G_2019_12, Sprint_3G_2019_12_map)
```

```{r}
T_Mobile_3G_2019_12 <- st_read(dsn = "F477_2019_12_Broadband_T_Mobile_80_280579", layer = "MBD_Technology86")
T_Mobile_4G_LTE_2019_12 <- st_read(dsn = "F477_2019_12_Broadband_T_Mobile_83_279545", layer = "MBD_LTE") 
T_Mobile_5G_2019_12 <- st_read(dsn = "F477_2019_12_Broadband_T_Mobile_84_280418", layer = "MBD_5G-NR")
```

```{r}
T_Mobile_4G_LTE_2019_12_map <- ggplot() +
  geom_sf(data = T_Mobile_4G_LTE_2019_12) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 4G LTE December 2019") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_4G_LTE_2019_12_map.pdf", plot = T_Mobile_4G_LTE_2019_12_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(T_Mobile_4G_LTE_2019_12, T_Mobile_4G_LTE_2019_12_map)
```


```{r}
T_Mobile_5G_2019_12_map <- ggplot() +
  geom_sf(data = T_Mobile_5G_2019_12) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 5G December 2019") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_5G_2019_12_map.pdf", plot = T_Mobile_5G_2019_12_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(T_Mobile_5G_2019_12_map, T_Mobile_5G_2019_12)
```

```{r}
T_Mobile_3G_2019_12_map <- ggplot() +
  geom_sf(data = T_Mobile_3G_2019_12) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 3G December 2019") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_3G_2019_12_map.pdf", plot = T_Mobile_3G_2019_12_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(T_Mobile_3G_2019_12, T_Mobile_3G_2019_12_map)
```

```{r}
Sprint_3G_2021_12 <- st_read(dsn = "F477_2021_12_Broadband_Sprint_85_384858", layer = "Sprint_EVDO_1")
Sprint_3G_2021_12_alternative <- st_read(dsn = "F477_2021_12_Broadband_Sprint_85_384862", layer = "Sprint_EVDO_2")
```
```{r}
Sprint_3G_2021_12_map <- ggplot() +
  geom_sf(data = Sprint_3G_2021_12) +
  geom_sf(data = Sprint_3G_2021_12_alternative) + 
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "Sprint 3G (EVDO 1) December 2021") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("Sprint_3G_2021_12_map.pdf", plot = Sprint_3G_2021_12_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(Sprint_3G_2021_12_map, Sprint_3G_2021_12, Sprint_3G_2021_12_alternative)
```



```{r}
T_Mobile_5G_2021_12_part_1 <- st_read(dsn = "F477_2021_12_Broadband_T_Mobile_89_384108", layer = "MBD_5G_NR_1")
T_Mobile_5G_2021_12_part_2 <- st_read(dsn = "F477_2021_12_Broadband_T_Mobile_89_384109", layer = "MBD_5G_NR_2")
```
```{r}
ggplot() +
  geom_sf(data = T_Mobile_5G_2021_12_part_1) +
    geom_sf(data = T_Mobile_5G_2021_12_part_2) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 5G December 2021") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_5G_2021_12_map.pdf", plot = T_Mobile_5G_2021_12_map, height = 4, width = 6, dpi = 72)
```

```{r}
rm(T_Mobile_5G_2021_12_part_1, T_Mobile_5G_2021_12_part_2, T_Mobile_5G_2019_12_map)
```

```{r}
T_Mobile_4G_2021_12_part_1 <- st_read(dsn = "F477_2021_12_Broadband_T_Mobile_83_383401", layer = "FCC_BDC_2")
T_Mobile_4G_2021_12_part_2 <- st_read(dsn = "F477_2021_12_Broadband_T_Mobile_83_399525", layer = "FCC_BDC_1")
```

```{r}
T_Mobile_4G_2021_12_map <- ggplot() +
  geom_sf(data = T_Mobile_4G_2021_12_part_1) +
    geom_sf(data = T_Mobile_4G_2021_12_part_1) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 4G December 2019") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_4G_2021_12_map.pdf", plot = T_Mobile_4G_2021_12_map, height = 4, width = 6, dpi = 72)
```


```{r}
T_Mobile_3G_2021_12_part_1 <- st_read(dsn = "F477_2021_12_Broadband_T_Mobile_86_390036", layer = "MBD_Technology86_1")
T_Mobile_3G_2021_12_part_2 <- st_read(dsn = "F477_2021_12_Broadband_T_Mobile_86_390037", layer = "MBD_Technology86_2")
```


```{r}
T_Mobile_3G_2021_12_map <- ggplot() +
  geom_sf(data = T_Mobile_3G_2021_12_part_1) +
    geom_sf(data = T_Mobile_3G_2021_12_part_2) +
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", linewidth = 0.1) +
  labs(title = "T-Mobile 3G December 2021") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        rect = element_blank())

ggsave("T_Mobile_3G_2021_12_map.pdf", plot = T_Mobile_3G_2021_12_map, height = 4, width = 6, dpi = 72)
```




